Detecting Hate Speech with GPT-3

### 2022-03-23 outputs/documents/paper.Rmd

---
output: 
  bookdown::pdf_document2:
    citation_package: natbib
    keep_tex: true
    toc: false
    fig_caption: true
    latex_engine: xelatex
    template: templates/svm-latex-ms.tex
bibliography: "references.bib"
header-includes:
  -  \usepackage{hyperref}
  -  \usepackage{amsmath}
biblio-style: apalike
title: "Detecting Hate Speech with GPT-3"
thanks: "Code and data are available at: https://github.com/kelichiu/GPT3-hate-speech-detection. We gratefully acknowledge the support of Gillian Hadfield, the Schwartz Reisman Institute for Technology and Society, and OpenAI for providing access to GPT-3 under the academic access program. We thank two anonymous reviews and the editor, as well as Amy Farrow, Haoluan Chen, John Giorgi, Mauricio Vargas Sepúlveda, Monica Alexander, Noam Kolt, and Tom Davidson for helpful discussions and suggestions. Please note that we have added asterisks to racial slurs and other offensive content in this paper, however the inputs and outputs did not have these.  Comments on the `r format(Sys.time(), '%d %B %Y')` version of this paper are welcome at: rohan.alexander@utoronto.ca."
author:
- name: Ke-Li Chiu
  affiliation: University of Toronto
- name: Annie Collins
  affiliation: University of Toronto
- name: Rohan Alexander
  affiliation: University of Toronto and Schwartz Reisman Institute
abstract: "Sophisticated language models such as OpenAI's GPT-3 can generate hateful text that targets marginalized groups. Given this capacity, we are interested in whether large language models can be used to identify hate speech and classify text as sexist or racist. We use GPT-3 to identify sexist and racist text passages with zero-, one-, and few-shot learning. We find that with zero- and one-shot learning, GPT-3 can identify sexist or racist text with an accuracy between 57 per cent and 68 per cent depending on the category of text and type of learning. With few-shot learning, the model's accuracy can be as high as 88 per cent. Large language models have a role to play in hate speech detection, and with further development could eventually be used to counter hate speech."
keywords: "GPT-3; natural language processing; quantitative analysis; hate speech."
date: "`r format(Sys.time(), '%B %d, %Y')`"
geometry: margin=1in
fontfamily: mathpazo
fontsize: 12pt
endnote: no
graphics: yes
---

@@

While these issues, among others, give us pause, the dual-use problem, which is that the same technology can be applied for both good and bad uses, provides motivation. For instance, while stylometric analysis can reveal the identity of political dissenters, it can also solve the unknown authorship of historic text [@hovy2016social]. In this paper we are interested in whether large language models, given that they can produce harmful language, can also identify, or learn to identify, harmful language.

Even though large NLP models do not have a real understanding of language, the vocabularies and the construction patterns of hateful language can be thought of as known to them. We show that this knowledge can be used to identify abusive language and even hate speech. We consider 120 different extracts that have been categorized as 'racist', 'sexist', or 'neither' in single-category settings (zero-shot, one-shot, and few-shot) and 243 different extracts in mixed-category few-shot settings. We ask GPT-3 to classify these based on zero-, one-, and few-shot learning, with and without instruction. We find that the model performs best with mixed-category few-shot learning. In that setting the model can accurately classify around 86 per cent of the racist extracts and 88 per cent of sexist extracts. If language models can be used to identify abusive language, then not only is there potential for them to counter the production of abusive language by humans, but they could also potentially self-police.

@@

We examine the ability of GPT-3 to identify hate speech in zero-shot, one-shot, and few-shot settings. There are a variety of parameters, such as temperature, that control the degree of text variation. Temperature is a hyper-parameter between zero and one. Lower temperatures mean that the model places more weight on higher-probability tokens. To explore potential variations in the classifications of comments, the temperature is set to 0.3 in our experiments for the zero- and one-shot settings and not varied. In all few-shot settings, the temperature is set to zero. There are two categories of hate speech that are of interest in this paper. The first targets the race of the recipient, and the other targets the gender of the recipient. With zero-, one-, and few-shot single-category learning, the model identifies hate speech one category at a time. With few-shot mixed-category learning, the categories are mixed, and the model is asked to classify an input as sexist, racist, or neither. Zero-shot learning means an example is not provided in the prompt. One-shot learning means that one example is provided, and few-shot means that two or more examples are provided.

@@

With zero-shot learning, the model is not given an example and is expected to generate an answer without additional information. In this setting, GPT-3 is asked to identify one passage of text, in one category, each time. A simple instruction is given asking the model to answer only 'Yes' or 'No'. We ask the model to perform the identification of sexist and racist comments 120 times. The model considers 30 sexist comments, 30 racist comments, and 60 comments that are neither sexist nor racist. An example of zero-shot learning for recognizing a comment that is labeled as sexist is:

@@

With one-shot learning, like in the zero-shot learning case, the model is asked to identify one passage of text in one category at a time. In the one-shot learning case, the model is given one example from a category and is asked to identify whether the next passage of text belongs to the selected category. Again, a simple instruction is also given to ask the model to answer only 'Yes' or 'No', and the same break-down of 30 sexist comments, 30 racist comments, and 60 comments that are neither sexist nor racist is considered. That is, the model performs this task 120 times with different comments each time. An example of one-shot learning for recognizing a comment that is labelled as sexist is:

@@

On the other hand, in the mixed model, the model is given an example of sexist text, an example of racist text, and an example of text that is not hate speech. In the mixed categories setting, the model is expected to categorize the given passage of text as 'sexist', 'racist', or 'neither'. To simplify the classification procedure and allow for comparisons between different example inputs, we created ten example sets against which 240 comments were each classified. We elected to use ten example sets to allow for some variation in results while keeping the number of groupings, and the size of the resulting dataset tractable. The example sets each consist of one sexist, one racist, and one 'neither' comment and were created using the comments at indexes one through ten from each dataset. Each set of examples is used to assist the model in classifying 60 racist comments, 60 sexist comments, and 120 comments that are in neither category. The inputted comments were the first _n_ + 1 comments from their respective datasets with the _i_^th^ comment dropped corresponding to the use of example set _i_. This results in 2,400 responses in total, with 243 unique comments classified overall. The ten sets of examples are shown in Appendix \@ref(appendxa). An example of this is:

@@

Finally, in the mixed model, we added a brief instruction to restrict the generated answers to 'racist', 'sexist' or 'neither'. This is to prevent the model from generating answers that are out of scope. An example of this is:

@@

For zero- and one-shot learning, each set of 120 comments was classified 100 times by GPT-3 in order to assess the variability of classifications at a temperature of 0.3. The reported performance metrics for these settings are the arithmetic means of each metric across all 100 iterations with the corresponding standard error. In some cases, the model outputted responses that were neither "yes" nor "no". These were counted as "not applicable" and omitted from analysis. Specific numbers are reported in the following sections.

@@

The overall results of the zero-shot experiments are presented in Table \@ref(tab:zeroshot-summary), and Appendix \@ref(appendixbzeroshot) provides additional detail. Out of 6000 classification for each category, the model has 3231 matches (true positives and negatives) and 2691 mismatches (false positives and negatives) in the sexist category, and 3463 matches and 2504 mismatches in the racist category. In the zero-shot learning setting, the model produced 111 responses that were inadmissible. The model performs more accurately when identifying racist comments, with an average accuracy of 58 per cent (SE = 6.5), compared with identifying sexist comments, with an average accuracy of 55 per cent (SE = 5.2). In contrast, the F1 score for classification of sexist speech is slightly higher on average at 63 per cent (SE = 4.3), compared with an average of 58 per cent (SE = 6.7) for racist speech. The overall ratio of matches and mismatches is 6694:5195. In other words, the average accuracy in identifying hate speech in the zero-shot setting is 56 per cent (SE = 4.6). The model has an average F1 score of 70 per cent (SE = 5.7) in this setting.

@@

The results of the one-shot learning experiments are presented in Table \@ref(tab:oneshot-summary), and Appendix \@ref(appendixboneshot) provides additional detail. The model has 37 matches and 23 mismatches in the racist category, and 36 matches and 24 mismatches in the sexist category. In contrast with the result generated from zero-shot learning, the model performs roughly the same in terms of identifying racist comments and identifying sexist comments. The overall ratio of matches and mismatches is 73:47. In other words, the accuracy of identifying hate speech in the one-shot setting is 61 per cent. The general performance in the one-shot setting is worse than in the zero-shot setting, with lower accuracy, recall, precision, and F1 score overall.

@@

The results of the single-category, few-shot learning, experiments are presented in Table \@ref(tab:fewshotsingle-summary), and Appendix \@ref(appendixbfewshotsingle) provides additional detail. The model has 41 matches and 19 mismatches in the racist category, and 42 matches and 18 mismatches in the sexist category. Like the one-shot setting, the model performs almost equally well in terms of identifying racist comments compared with identifying sexist comments. The general performance in the single-category few-shot learning setting is slightly more accurate than performance in other settings, with an accuracy of 69 per cent compared with 61 per cent in the one-shot setting and 63 per cent in the zero-shot setting. But, in this context we see the lowest F1 score of the settings considered thus far at 61 per cent.

```{r fewshotsingle-summary, message = FALSE}
# Set-up data for zero-shot racism
few_single_shot_result <-
  read_csv(here::here("outputs/data/few_shot_single_results_multiple.csv"),
           show_col_types = FALSE)

@@

few_shot_result_hate <- 
    few_single_shot_result %>%
    select(category, label, answer, temperature) %>%
    mutate(label_yn = ifelse(grepl("not", label), "N", "Y"),
           answer_yn = ifelse(grepl("not", label), "N", "Y")) %>%
    count(label_yn, answer_yn) %>%
    pivot_wider(names_from = answer_yn, values_from = n, values_fill = 0) %>%
    mutate(label_yn = case_when(label_yn == "N" ~ "Not hate speech",
                                label_yn == "Y" ~ "Hate speech")) %>%
    rename("Actual classification" = label_yn, "Not hate speech" = N, "Hate speech" = Y)
few_shot_result_hate <- few_shot_result_hate[, c("Actual classification", "Not hate speech", "Hate speech")]

@@

few_single_summary %>% pivot_longer(cols = !c(name, calc), names_to = "metric") %>%
  pivot_wider(names_from = c(calc), values_from = value) %>%
  mutate(name = case_when(metric == "accuracy" ~ name, TRUE ~ ""),
         metric = str_to_title(metric),
         Mean = round(Mean),
         SE = round(SE, 1)) %>%
  knitr::kable(col.names = c("", "Metric", "Mean (%)", "Standard Error (%)"), 
        caption = "Performance of model in zero-shot learning across 100 classifications of each comment at a temperature of 0.3.",
        booktabs = TRUE,
        linesep = "") %>%
  kableExtra::column_spec(column = 1, bold = TRUE)
```

@@

The results of the mixed-category few-shot experiments are presented in Table \@ref(tab:fewshotmixed-summary), and Appendix \@ref(appendxbmuxedubsnorinstruction) provides additional detail. Among the ten sets of examples, Example Set 1 yields the best performance in terms of accuracy (91 per cent) and F1 score (87 per cent) for racist comments and Example Set 10 yields the best performance by the same metrics for sexist comments (accuracy of 89 per cent, F1 score of 85 per cent). The example set that yields the worst results in identifying racist text in terms of F1 score is Example Set 8, which has an F1 score of 70 per cent (and the lowest accuracy at 73 per cent) for this dataset. The example set that yields the worst results in identifying sexist text in terms of F1 score is Example Set 9, which has an F1 score of 69 per cent (and the lowest accuracy at 74 per cent) for this dataset. The differences between Example Sets 1, 8, 9, and 10 suggest that, although the models are provided with the same number of examples, the content of the examples also affects how the model makes inferences. Overall, the mixed-category few-shot setting performs roughly the same in terms of identifying sexist text and racist text. It also has higher accuracy and F1 score than the zero-shot, one-shot, and single-category few-shot settings for both racist and sexist text.

```{r fewshotmixed-summary}
# Racism
few_shot_mixed_result <-
  read.csv(here::here("outputs/data/few_shot_fixed_examples_results.csv"))

# Example sets
few_shot_mixed_result_racism <- 
  few_shot_mixed_result %>%
  filter(label != "sexist") %>%
  mutate(example_set = example_set + 1,
         label = case_when(!grepl("racist", label) ~ "not racist", TRUE ~ "racist"),
         answer = case_when(!grepl("racist", answer) ~ "not racist", TRUE ~ "racist")) %>%
  count(example_set, label, answer) %>%
  pivot_wider(names_from = answer, values_from = n) %>%
  mutate(label = case_when(label == "not racist" ~ "Not racist",
                           label == "racist" ~ "Racist")) %>%
  rename("Actual classification" = label, "Not racist" = "not racist", "Racist" = "racist")

# Overall
few_shot_mixed_result_racism_overall <- 
  few_shot_mixed_result %>%
  filter(label != "sexist") %>%
  mutate(example_set = example_set + 1,
         label = case_when(!grepl("racist", label) ~ "not racist", TRUE ~ "racist"),
         answer = case_when(!grepl("racist", answer) ~ "not racist", TRUE ~ "racist")) %>%
  count(label, answer) %>%
  pivot_wider(names_from = answer, values_from = n) %>%
  mutate(label = case_when(label == "not racist" ~ "Not racist",
                           label == "racist" ~ "Racist")) %>%
  rename("Actual classification" = label, "Not racist" = "not racist", "Racist" = "racist")

@@

# Sexism
# Example sets
few_shot_mixed_result_sexism <- 
  few_shot_mixed_result %>%
  filter(label != "racist") %>%
  mutate(example_set = example_set + 1,
         label = case_when(!grepl("sexist", label) ~ "not sexist", TRUE ~ "sexist"),
         answer = case_when(!grepl("sexist", answer) ~ "not sexist", TRUE ~ "sexist")) %>%
  count(example_set, label, answer) %>%
  pivot_wider(names_from = answer, values_from = n) %>%
  mutate(label = case_when(label == "not sexist" ~ "Not sexist",
                           label == "sexist" ~ "Sexist")) %>%
  rename("Actual classification" = label, "Not sexist" = "not sexist", "Sexist" = "sexist")

# Overall
few_shot_mixed_result_sexism_overall <- 
  few_shot_mixed_result %>%
  filter(label != "racist") %>%
  mutate(example_set = example_set + 1,
         label = case_when(!grepl("sexist", label) ~ "not sexist", TRUE ~ "sexist"),
         answer = case_when(!grepl("sexist", answer) ~ "not sexist", TRUE ~ "sexist")) %>%
  count(label, answer) %>%
  pivot_wider(names_from = answer, values_from = n) %>%
  mutate(label = case_when(label == "not sexist" ~ "Not sexist",
                           label == "sexist" ~ "Sexist")) %>%
  rename("Actual classification" = label, "Not sexist" = "not sexist", "Sexist" = "sexist")

@@

@@

The unique generated answers are listed in Table \@ref(tab:fewshotmixedanswersnoinstruct). These are the response of GPT-3 that we obtain when we ask the model to classify statements, but do not provide examples that would serve to limit the responses. Under the mixed-category setting, the model generates many answers that are out of scope. For instance, other than 'sexist', 'racist', and 'neither', we also see answers such as 'transphobic', 'hypocritical', 'Islamophobic', and 'ableist'. In some cases, the model even classifies a text passage into more than one category, such as 'sexist, racist' and 'sexist, misogynistic'. The full list contains eighty different answers instead of three.

@@

```{r fewshotmixedanswersnoinstruct, warning = FALSE, message = FALSE}
few_shot_mixed_result <-
  readr::read_csv(here::here("outputs/data/few_shot_fixed_examples_results.csv"))

@@

To reduce the chance of the model generating answers that are out of scope, a brief instruction is added to the prompt, specifying that the answers be: 'sexist', 'racist', or 'neither'. The addition of an instruction successfully restricts the generated answers within the specified terms. The unique generated answers are: 'racist', 'sexist', and 'neither'.

The results of the mixed-category few-shot learning, with instruction, experiments are presented in Tables \@ref(tab:fewshotmixedinstruct-matrix) and \@ref(tab:fewshotmixedinstruct-summary), and Appendix \@ref(appendxbmuxedubstryctub) provides additional detail. With the addition of an instruction in the prompt, Example Set 10 remains the best performing example set in terms of accuracy (89 per cent) and F1 score (84 per cent) for sexist text, however Example Set 4 now performs best for racist text in terms of both accuracy (89 per cent) and F1 score (85 per cent). When considering the classification of racist and sexist speech overall, the models perform similarly with and without instruction when classifying racist text, but the model appears to perform slightly better at identifying sexist text when the instruction is omitted.

However, examining label-classification matches across all categories ('sexist', 'racist', and 'neither'), mixed-category few-shot learning almost always performs better with instruction than without instruction (Figure \ref{fig:comparison}). Across all example sets, the mean proportion of matching classifications (out of 240 comments) for mixed-category few-shot learning without instruction is 67 per cent. The average proportion of matching classifications rises to 75 per cent for learning with instruction.

```{r fewshotmixedinstruct-matrix}
few_shot_instruction_result <-
  read.csv(here::here(
    "outputs/data/few_shot_fixed_examples_instruction_results.csv"
  ))

confusion_instruct <- few_shot_instruction_result %>%
  count(label, answer) %>%
  pivot_wider(names_from = answer, values_from = n, values_fill = 0) %>%
  mutate(label = case_when(label == "neither" ~ "Neither",
                           label == "racist" ~ "Racist",
                           label == "sexist" ~ "Sexist")) %>%
  rename("Actual classification" = label, Neither = " neither", Racist = " racist", Sexist = " sexist")

>>>>

---
output: 
  bookdown::pdf_document2:
    citation_package: natbib
    keep_tex: true
    toc: false
    fig_caption: true
    latex_engine: xelatex
    template: templates/svm-latex-ms.tex
bibliography: "references.bib"
header-includes:
  -  \usepackage{hyperref}
  -  \usepackage{amsmath}
biblio-style: apalike
title: "Detecting Hate Speech with GPT-3"
thanks: "Code and data are available at: https://github.com/kelichiu/GPT3-hate-speech-detection. We gratefully acknowledge the support of Gillian Hadfield, the Schwartz Reisman Institute for Technology and Society, and OpenAI for providing access to GPT-3 under the academic access program. We thank two anonymous reviews and the editor, as well as Amy Farrow, Haoluan Chen, John Giorgi, Mauricio Vargas Sepúlveda, Monica Alexander, Noam Kolt, and Tom Davidson for helpful discussions and suggestions. Please note that we have added asterisks to racial slurs and other offensive content in this paper, however the inputs and outputs did not have these.  Comments on the `r format(Sys.time(), '%d %B %Y')` version of this paper are welcome at: rohan.alexander@utoronto.ca."
author:
- name: Ke-Li Chiu
  affiliation: University of Toronto
- name: Annie Collins
  affiliation: University of Toronto
- name: Rohan Alexander
  affiliation: University of Toronto and Schwartz Reisman Institute
abstract: "Sophisticated language models such as OpenAI's GPT-3 can generate hateful text that targets marginalized groups. Given this capacity, we are interested in whether large language models can be used to identify hate speech and classify text as sexist or racist. We use GPT-3 to identify sexist and racist text passages with zero-, one-, and few-shot learning. We find that with zero- and one-shot learning, GPT-3 can identify sexist or racist text with an average accuracy between 55 per cent and 67 per cent depending on the category of text and type of learning. With few-shot learning, the model's accuracy can be as high as 85 per cent. Large language models have a role to play in hate speech detection, and with further development could eventually be used to counter hate speech."
keywords: "GPT-3; natural language processing; quantitative analysis; hate speech."
date: "`r format(Sys.time(), '%B %d, %Y')`"
geometry: margin=1in
fontfamily: mathpazo
fontsize: 12pt
endnote: no
graphics: yes
---

@@

While these issues, among others, give us pause, the dual-use problem, which is that the same technology can be applied for both good and bad uses, provides motivation. For instance, while stylometric analysis can reveal the identity of political dissenters, it can also solve the unknown authorship of historic text [@hovy2016social]. In this paper we are interested in whether large language models, given that they can produce harmful language, can also identify (or learn to identify) harmful language.

Even though large NLP models do not have a real understanding of language, the vocabularies and the construction patterns of hateful language can be thought of as known to them. We show that this knowledge can be used to identify abusive language and even hate speech. We consider 120 different extracts that have been categorized as 'racist', 'sexist', or 'neither' in single-category settings (zero-shot, one-shot, and few-shot) and 243 different extracts in mixed-category few-shot settings. We ask GPT-3 to classify these based on zero-, one-, and few-shot learning, with and without instruction. We find that the model performs best with mixed-category few-shot learning. In that setting the model can accurately classify around 83 per cent of the racist extracts and 85 per cent of sexist extracts on average, with F1 scores of 79 per cent and 77 per cent respectively. If language models can be used to identify abusive language, then not only is there potential for them to counter the production of abusive language by humans, but they could also potentially self-police.

@@

<!-- Fix once few-shot results are all in -->

We examine the ability of GPT-3 to identify hate speech in zero-shot, one-shot, and few-shot settings. There are a variety of parameters, such as temperature, that control the degree of text variation. Temperature is a hyper-parameter between zero and one. Lower temperatures mean that the model places more weight on higher-probability tokens. To explore the variability in the classifications of comments, the temperature is set to 0.3 in our experiments. There are two categories of hate speech that are of interest in this paper. The first targets the race of the recipient, and the other targets the gender of the recipient. With zero-, one-, and few-shot single-category learning, the model identifies hate speech one category at a time. With few-shot mixed-category learning, the categories are mixed, and the model is asked to classify an input as sexist, racist, or neither. Zero-shot learning means an example is not provided in the prompt. One-shot learning means that one example is provided, and few-shot means that two or more examples are provided. All classification tasks were performed on the Davinci engine, GPT-3's most powerful and recently trained engine.

@@

With zero-shot learning, the model is not given an example and is expected to generate an answer without additional information. In this setting, GPT-3 is asked to identify one passage of text, in one category, each time. A simple instruction is given asking the model to answer only 'Yes' or 'No'. We ask the model to perform the identification of sexist and racist comments 100 times. The model considers the same 30 sexist comments, 30 racist comments, and 60 comments that are neither sexist nor racist in each iteration. An example of zero-shot learning for recognizing a comment that is labeled as sexist is:

@@

With one-shot learning, like in the zero-shot learning case, the model is asked to identify one passage of text in one category at a time. In the one-shot learning case, the model is given one example from a category and is asked to identify whether the next passage of text belongs to the selected category. Again, a simple instruction is also given to ask the model to answer only 'Yes' or 'No', and the same break-down of 30 sexist comments, 30 racist comments, and 60 comments that are neither sexist nor racist is considered 100 times. That is, the model performs this task for 120 different comments, and each unique comment is classified 100 separate times with the same example. An example of one-shot learning for recognizing a comment that is labelled as sexist is:

@@

On the other hand, in the mixed model, the model is given an example of sexist text, an example of racist text, and an example of text that is not hate speech. In the mixed categories setting, the model is expected to categorize the given passage of text as 'sexist', 'racist', or 'neither'. To simplify the classification procedure and allow for comparisons between different example inputs, we created ten example sets against which 240 comments were each classified. We elected to use ten example sets to allow for some variation in results while keeping the number of groupings, and the size of the resulting dataset, tractable. The example sets each consist of one sexist, one racist, and one 'neither' comment and were created using the comments at indexes one through ten from each dataset. Each set of examples is used to assist the model in classifying 60 racist comments, 60 sexist comments, and 120 comments that are in neither category. The inputted comments were the first _n_ + 1 comments from their respective datasets with the _i_^th^ comment dropped corresponding to the use of example set _i_. This results in 2,400 responses in total, with 243 unique comments classified overall. The ten sets of examples are shown in Appendix \@ref(appendxa). An example of this is:

@@

Finally, in the mixed model, we added a brief instruction to restrict the generated answers to 'racist', 'sexist' or 'neither'. This is an attempt to prevent the model from generating answers that are out of scope. An example of this is:

@@

For zero- and one-shot learning, each set of 120 comments was classified 100 times by GPT-3 in order to assess the variability of classifications at a temperature of 0.3. The reported performance metrics for these settings are the arithmetic means of each metric across all 100 iterations with the corresponding standard error. In the zero-shot setting, the model sometimes outputted responses that were neither "yes" nor "no". These were counted as "not applicable" and omitted from analysis. Specific numbers are reported in the following section.

@@

The overall results of the zero-shot experiments are presented in Table \@ref(tab:zeroshot-summary), and Appendix \@ref(appendixbzeroshot) provides additional detail. Out of 6000 classification for each category, the model has 3231 matches (true positives and negatives) and 2691 mismatches (false positives and negatives) in the sexist category, and 3463 matches and 2504 mismatches in the racist category. In this setting, the model sometimes outputted responses that were neither "yes" nor "no". This occurred for 111 classifications, which were subsequently omitted from analysis. The model performs more accurately when identifying racist comments, with an average accuracy of 58 per cent (SE = 6.5), compared with identifying sexist comments, with an average accuracy of 55 per cent (SE = 5.2). In contrast, the F1 score for classification of sexist speech is slightly higher on average at 63 per cent (SE = 4.3), compared with an average of 58 per cent (SE = 6.7) for racist speech. The overall ratio of matches and mismatches is 6694:5195. In other words, the average accuracy in identifying hate speech in the zero-shot setting is 56 per cent (SE = 4.6). The model has an average F1 score of 70 per cent (SE = 5.7) in this setting.

@@

The results of the one-shot learning experiments are presented in Table \@ref(tab:oneshot-summary), and Appendix \@ref(appendixboneshot) provides additional detail. Out of 6000 classifications each, the model produced 3284 matches and 2668 mismatches in the racist category, and 3236 matches and 2631 mismatches in the sexist category. Unlike the results generated from zero-shot learning, the model performs approximately the same when identifying sexist and racist comments, with an average accuracy of 55 per cent (SE = 6.4) and an F1 score of 58 per cent (SE = 6.5) when identifying racist comments, compared with sexist comments at an accuracy of 55 per cent (SE = 5.8) and an F1 score of 56 per cent (SE = 6.3). The overall ratio of matches and mismatches is 6520:5326. In other words, the average accuracy of identifying hate speech in the one-shot setting is 55 per cent (SE = 4.1). The general performance in the one-shot setting is nearly the same as in the zero-shot setting, with an overall average accuracy of 55 per cent compared with 56 per cent (SE = 4.6) in the zero-shot setting. However, the F1 score in the one-shot setting is much lower than in the zero-shot setting at 55 per cent (SE = 7.3) compared with 70 per cent (SE = 5.7).

@@

The results of the single-category, few-shot learning, experiments are presented in Table \@ref(tab:fewshotsingle-summary), and Appendix \@ref(appendixbfewshotsingle) provides additional detail. The model has 3862 matches and 2138 mismatches in the racist category, and 4209 matches and 1791 mismatches in the sexist category. Unlike in the zero- and one-shot settings, the model performs slightly better when identifying sexist comments compared with identifying racist comments. The general performance in the single-category few-shot learning setting is more accurate than performance in other settings, with an accuracy of 67 per cent (SE = 2.7) compared with 55 per cent in the one-shot setting (SE = 4.1) and 56 per cent (SE = 4.3) in the zero-shot setting. The average F1 score in this setting is 62 per cent (SE = 4.9) which is similar to the results of the one-shot setting but slightly lower than in the zero-shot setting.

```{r fewshotsingle-summary, message = FALSE}
# Set-up data for few-shot racism
few_single_shot_result <-
  read_csv(here::here("outputs/data/few_shot_single_results_multiple.csv"),
           show_col_types = FALSE)

@@

few_shot_result_hate <- 
    few_single_shot_result %>%
    select(category, label, answer, temperature) %>%
    mutate(label_yn = ifelse(grepl("not", label), "N", "Y"),
           answer_yn = ifelse(grepl("not", answer), "N", "Y")) %>%
    count(label_yn, answer_yn) %>%
    pivot_wider(names_from = answer_yn, values_from = n, values_fill = 0) %>%
    mutate(label_yn = case_when(label_yn == "N" ~ "Not hate speech",
                                label_yn == "Y" ~ "Hate speech")) %>%
    rename("Actual classification" = label_yn, "Not hate speech" = N, "Hate speech" = Y)
few_shot_result_hate <- few_shot_result_hate[, c("Actual classification", "Not hate speech", "Hate speech")]

@@

few_single_summary %>% pivot_longer(cols = !c(name, calc), names_to = "metric") %>%
  pivot_wider(names_from = c(calc), values_from = value) %>%
  mutate(name = case_when(metric == "accuracy" ~ name, TRUE ~ ""),
         metric = str_to_title(metric),
         Mean = round(Mean),
         SE = round(SE, 1)) %>%
  knitr::kable(col.names = c("", "Metric", "Mean (%)", "Standard Error (%)"), 
        caption = "Performance of model in single category few-shot learning across 100 classifications of each comment at a temperature of 0.3.",
        booktabs = TRUE,
        linesep = "") %>%
  kableExtra::column_spec(column = 1, bold = TRUE)
```

@@

The results of the mixed-category few-shot experiments are presented in Table \@ref(tab:fewshotmixed-summary), and Appendix \@ref(appendxbmuxedubsnorinstruction) provides additional detail. Among the ten sets of examples, Example Set 10 yields the best performance in terms of accuracy (91 per cent) and F1 score (87 per cent) for racist comments. The model performs with similar accuracy for identifying racist comments across most of the example sets (approximately 87 per cent), however the highest F1 score results from Example Set 10 once again. The example set that yields the worst results in identifying racist text in terms of F1 score is Example Set 8, which has an F1 score of 69 per cent (and the lowest accuracy at 70 per cent) for this dataset. The example set that yields the worst results in identifying sexist text in terms of F1 score is Example Set 9, which has an F1 score of 69 per cent (and the lowest accuracy at 76 per cent) for this dataset. The differences between Example Sets 8, 9, and 10 suggest that, although the models are provided with the same number of examples, the content of the examples also affects how the model makes inferences. Overall, the mixed-category few-shot setting performs roughly the same in terms of identifying sexist text and racist text. It also has distinctly higher accuracy and F1 score overall than the zero-shot, one-shot, and single-category few-shot settings for both racist and sexist text.

```{r fewshotmixed-summary}
# Racism
few_shot_mixed_result <-
  read.csv(here::here("outputs/data/few_shot_fixed_examples_results_multiple.csv"))

# Keep only first iteration
few_shot_mixed_result <- few_shot_mixed_result %>% slice(2401:4800)

# Example sets
few_shot_mixed_result_racism <- 
  few_shot_mixed_result %>%
  filter(label != "sexist") %>%
  mutate(example_set = example_set + 1,
         label = case_when(!grepl("racist", label) ~ "not racist", TRUE ~ "racist"),
         answer = case_when(!grepl("racist", answer) ~ "not racist", TRUE ~ "racist")) %>%
  count(example_set, label, answer) %>%
  pivot_wider(names_from = answer, values_from = n, values_fill = 0) %>%
  mutate(label = case_when(label == "not racist" ~ "Not racist",
                           label == "racist" ~ "Racist")) %>%
  rename("Actual classification" = label, "Not racist" = "not racist", "Racist" = "racist")

# Overall
few_shot_mixed_result_racism_overall <- 
  few_shot_mixed_result %>%
  filter(label != "sexist") %>%
  mutate(example_set = example_set + 1,
         label = case_when(!grepl("racist", label) ~ "not racist", TRUE ~ "racist"),
         answer = case_when(!grepl("racist", answer) ~ "not racist", TRUE ~ "racist")) %>%
  count(label, answer) %>%
  pivot_wider(names_from = answer, values_from = n, values_fill = 0) %>%
  mutate(label = case_when(label == "not racist" ~ "Not racist",
                           label == "racist" ~ "Racist")) %>%
  rename("Actual classification" = label, "Not racist" = "not racist", "Racist" = "racist")

@@

# Sexism
# Example sets
few_shot_mixed_result_sexism <- 
  few_shot_mixed_result %>%
  filter(label != "racist") %>%
  mutate(example_set = example_set + 1,
         label = case_when(!grepl("sexist", label) ~ "not sexist", TRUE ~ "sexist"),
         answer = case_when(!grepl("sexist", answer) ~ "not sexist", TRUE ~ "sexist")) %>%
  count(example_set, label, answer) %>%
  pivot_wider(names_from = answer, values_from = n, values_fill = 0) %>%
  mutate(label = case_when(label == "not sexist" ~ "Not sexist",
                           label == "sexist" ~ "Sexist")) %>%
  rename("Actual classification" = label, "Not sexist" = "not sexist", "Sexist" = "sexist")

# Overall
few_shot_mixed_result_sexism_overall <- 
  few_shot_mixed_result %>%
  filter(label != "racist") %>%
  mutate(example_set = example_set + 1,
         label = case_when(!grepl("sexist", label) ~ "not sexist", TRUE ~ "sexist"),
         answer = case_when(!grepl("sexist", answer) ~ "not sexist", TRUE ~ "sexist")) %>%
  count(label, answer) %>%
  pivot_wider(names_from = answer, values_from = n, values_fill = 0) %>%
  mutate(label = case_when(label == "not sexist" ~ "Not sexist",
                           label == "sexist" ~ "Sexist")) %>%
  rename("Actual classification" = label, "Not sexist" = "not sexist", "Sexist" = "sexist")

@@

  few_shot_summary <- tibble(example_set = rep(i, 4),
                             Category = rep("Racism", 4), 
                             Metric = c("Accuracy", "Precision", "Recall", "F1"), 
                             Percent = few_racism_all) %>%
  pivot_wider(names_from = Metric, values_from = Percent, values_fill = 0)

@@

The unique generated answers are listed in Table \@ref(tab:fewshotmixedanswersnoinstruct). These are the response of GPT-3 that we obtain when we ask the model to classify statements, but do not provide examples that would serve to limit the responses. Under the mixed-category setting, the model generates many answers that are out of scope. For instance, other than 'sexist', 'racist', and 'neither', we also see answers such as 'transphobic', 'hypocritical', 'Islamophobic', and 'ableist'. In some cases, the model even classifies a text passage into more than one category, such as 'sexist, racist' and 'sexist and misogynistic'. The full list contains 143 different answers instead of three.

@@

```{r fewshotmixedanswersnoinstruct, warning = FALSE, message = FALSE}
few_shot_mixed_result <-
  read.csv(here::here("outputs/data/few_shot_fixed_examples_results_multiple.csv"))

# Keep only first iteration
few_shot_mixed_result <- few_shot_mixed_result %>% slice(2401:4800)

@@

To reduce the chance of the model generating answers that are out of scope, a brief instruction is added to the prompt, specifying that the answers be: 'sexist', 'racist', or 'neither'. The addition of an instruction successfully restricts the generated answers within the specified terms with the exception of three responses: one classification of "racist and sexist" and two classifications of "both". These responses were likely a result of randomness introduced by the non-zero temperature and were excluded from analysis. The unique generated answers are: 'racist', 'sexist', 'neither', 'both', and 'racist and sexist'.

The results of the mixed-category few-shot learning, with instruction, experiments are presented in Tables \@ref(tab:fewshotmixedinstruct-matrix) and \@ref(tab:fewshotmixedinstruct-summary), and Appendix \@ref(appendxbmuxedubstryctub) provides additional detail. With the addition of an instruction in the prompt, Example Set 10 remains the best performing example set in terms of accuracy (86 per cent) and F1 score (78 per cent) for sexist text. Performance in classifying racist text is slightly more varied in this setting, with Example Set 7 performing most accurately at 88 per cent (and with the highest F1 score at 82 per cent). Considering the classification of racist and sexist speech overall, the models perform similarly with and without instruction when classifying racist text, but the model appears to perform slightly better at identifying sexist text when the instruction is omitted.

However, examining label-classification matches across all categories ('sexist', 'racist', and 'neither'), mixed-category few-shot learning almost always performs better with instruction than without instruction (Figure \ref{fig:comparison}). Across all example sets, the mean proportion of matching classifications (out of 240 comments) for mixed-category few-shot learning without instruction is 65 per cent. The average proportion of matching classifications rises to 71 per cent for learning with instruction.

```{r fewshotmixedinstruct-matrix}
few_shot_instruction_result <-
  read.csv(here::here(
    "outputs/data/few_shot_fixed_example_instruction_results.csv"
  ))

confusion_instruct <- few_shot_instruction_result %>%
  count(label, answer) %>%
  mutate(answer = str_to_title(answer) %>% trimws(),
         label = str_to_title(label) %>% trimws()) %>%
  pivot_wider(names_from = answer, values_from = n, values_fill = 0) %>%
  rename(`Actual classification` = label)
